The very dependent recursive structure of iterated parametricity in indexed form

### 2025-12-23 paper/paper.tex
paper: better typesetting with some error fixes

The types $\paintingtype[n][\lbrack 0,p-1 \rbrack]$ of $\painting[n][\lbrack 0,p \rbrack]$ for $p \leq n+1$ are type families over frames, so $\framep[n][\lbrack 0,p-1 \rbrack]$ needs to be assumed beforehand:

@@

    \multicolumn{3}{l}{
    (\deps[n][\lbrack 0,p-1 \rbrack]:\depstype[n][\lbrack 0,p-1 \rbrack]) :\frametype[n+1][\lbrack 0,p \rbrack]}         \\

@@

    \unitpoint                            & \defeq & (\unitpoint,unit)                                                   \\

@@

    \multicolumn{3}{l}{

@@

            \end{array}\right) : \U} \\

@@

    : \U}                                                                                                                \\

@@

      : \layertype[n][p]
      (\deps[n][\lbrack 0,p \rbrack])}

@@

Let us introduce another abbreviation to combine a full $\deps[n][\lbrack 0,n \rbrack]:\depstype[n][\lbrack 0,n \rbrack]$ with a family $E:\framep[n+1][n+1](\deps[n][\lbrack 0,n \rbrack])\rightarrow \U$:
\begin{gather*}
  \fulldepstype[n] \defeq
  \left(\begin{array}{lll}
      \deps[n][\lbrack 0,n \rbrack] & : & \depstype[n][\lbrack 0,n \rbrack]                               \\
      E                             & : & \framep[n+1][n+1](\deps[n][\lbrack 0,n \rbrack]) \rightarrow \U \\
    \end{array}\right)
\end{gather*}

@@

\begin{equation*}
  \begin{array}{llcl}
    \restrptype[n][\lbrack 0,p-1 \rbrack] &
    (\fulldeps[n]:\fulldepstype[n])       & :      & \U                                                                \\
    \restrptype[n]                        & E      & \defeq                                                & \unittype \\
    \restrptype[n][\lbrack 0,p \rbrack]   &
    (\fulldeps[n])
                                          & \defeq & \restrptype[n+1][\lbrack 0,p-1 \rbrack] \times \left(
    \begin{array}{l}
        \Pi q \leq n-p.\Pi \epsilon.\Pi d:\framep[n+1][p](\fulldeps.\deps[n][\lbrack 0,p-1 \rbrack]). \\
        \painting[n+1][p]
        (\fulldeps[n])(d) \rightarrow                                                                 \\ \fulldeps.\deps.\painting[n][p](\fulldeps.\deps.\restrf[n][p][q][\epsilon](d))\end{array}
    \right)                                                                                                            \\
  \end{array}
\end{equation*}

@@

    \multicolumn{3}{l}{
    (\depscohs[n][\lbrack 0,p-1 \rbrack]:\depscohstype[n][\lbrack 0,p-1 \rbrack]) : \depstype[n+1][\lbrack 0,p \rbrack]}                                                                                \\

@@

    \multicolumn{3}{l}{
      (\depscohs[n][\lbrack 0,p-1 \rbrack]:\depscohstype[n][\lbrack 0,p-1 \rbrack]) :

@@

              \end{array}\right)
    }                                                                                                                                                                                                   \\

@@

            \end{array}\right) : \U
    }                                                                                                                  n                                                                                \\

@@

                               \end{array}\right)          . \\

@@

    \multicolumn{3}{l}{
      (\depscohs[n][\lbrack 0,p \rbrack]:\depscohstype[n][\lbrack 0,p \rbrack]) :
    \restrltype[n][p] (\depscohs[n][\lbrack 0,p \rbrack])}                                                                                                                                              \\

@@

We can then extend $\overline{\deps}$ to build $\fulldeps$ from $\fulldepscohs$:
\begin{align*}
  \overline{\fulldeps}(\fulldepscohs[n]:\fulldepscohstype[n]) & :      & \fulldepstype[n+1] \\
  \overline{\fulldeps}(\fulldepscohs[n])                      & \defeq &
  \left(\begin{array}{l}
            \overline{\deps}(\fulldepscohs.\depscohs[n][\lbrack 0,n \rbrack]) \\
            \fulldepscohs.E                                                   \\
          \end{array}\right)                   \\
\end{align*}

@@

\begin{align*}{4}
  \restrp[n+1][\lbrack 0,p-1 \rbrack]     &
  (\fulldepscohs[n]:\fulldepscohstype[n]) & :      & \restrptype[n+1][\lbrack 0,p-1 \rbrack](\fulldepscohs.\depscohs.\fulldeps[n])) \\
  \restrp[n+1]                            &
  \fulldepscohs[n]                        & \defeq & \unittype                                                                      \\
  \restrp[n+1][\lbrack 0,p \rbrack]       &
  \fulldepscohs[n]
                                          & \defeq &
  (\restrp[n+1][\lbrack 0,p-1 \rbrack](\fulldepscohs[n]),\restrp[n+1][p](\fulldepscohs[n]))                                         \\
\end{align*}

@@

\begin{align*}
  \overline{\fulldepscohs}(\fulldepscoht[n]:\fulldepscohstype[n]) & :      & \fulldepscohstype[n+1] \\
  \overline{\fulldepscohs}(\fulldepscoht[n])                      & \defeq &
  \left(\begin{array}{l}
            \overline{\depscohs}(\fulldepscoht.\depscoht[n][\lbrack 0,n \rbrack]) \\
            \fulldepscoht.E                                                       \\
          \end{array}\right)                       \\
\end{align*}

>>>>

The types $\paintingtype[n][\lbrack 0,p-1 \rbrack]$ of $\painting[n][\lbrack 0,p-1 \rbrack]$ for $p \leq n+1$ are type families over frames, so $\framep[n][\lbrack 0,p-1 \rbrack]$ needs to be assumed beforehand:

@@

    (\deps[n][\lbrack 0,p-1 \rbrack]:\depstype[n][\lbrack 0,p-1 \rbrack])                                             & :      & \frametype[n+1][\lbrack 0,p \rbrack]   \\

@@

    \unitpoint                                                                                                        & \defeq & (\unitpoint,\unittype)                 \\

@@

@@

            \end{array}\right) & :      & \U                                                \\

@@

                                                                                                                      & :      & \U                                     \\

@@

                                                                                                                      & :      & \layertype[n][p]

@@

Let us introduce another abbreviation to combine a full $\deps[n][\lbrack 0,n \rbrack]:\depstype[n][\lbrack 0,n \rbrack]$ with a family $E:\framep[n+1][n+1](\deps[n][\lbrack 0,n \rbrack])\rightarrow \U$:
\begin{equation*}
  \fulldepstype[n] \defeq
  \left(\begin{array}{lcl}
      \deps[n][\lbrack 0,n \rbrack] & : & \depstype[n][\lbrack 0,n \rbrack]                               \\
      E                             & : & \framep[n+1][n+1](\deps[n][\lbrack 0,n \rbrack]) \rightarrow \U \\
    \end{array}\right)
\end{equation*}

@@

\begin{equation*}
  \begin{array}{llcl}
    \restrptype[n][\lbrack 0,p-1 \rbrack] &
    (\fulldeps[n]:\fulldepstype[n])       & :      & \U                                                         \\
    \restrptype[n]                        & E      & \defeq                                         & \unittype \\
    \restrptype[n][\lbrack 0,p \rbrack]   &
    (\fulldeps[n])
                                          & \defeq & \restrptype[n+1][\lbrack 0,p-1 \rbrack] \times             \\
                                          &        &                                                & \left(
    \begin{array}{l}
        \Pi q \leq n-p.\Pi \epsilon.\Pi d:\framep[n+1][p](\fulldeps.\deps[n][\lbrack 0,p-1 \rbrack]). \\
        \painting[n+1][p]
        (\fulldeps[n])(d) \rightarrow                                                                 \\ \fulldeps.\deps.\painting[n][p](\fulldeps.\deps.\restrf[n][p][q][\epsilon](d))\end{array}
    \right)                                                                                                     \\
  \end{array}
\end{equation*}

@@

    (\depscohs[n][\lbrack 0,p-1 \rbrack]:\depscohstype[n][\lbrack 0,p-1 \rbrack])            & :                                   & \depstype[n+1][\lbrack 0,p \rbrack]                                \\

@@

    (\depscohs[n][\lbrack 0,p-1 \rbrack]:\depscohstype[n][\lbrack 0,p-1 \rbrack])            & :                                   &

@@

@@

            \end{array}\right) & :                                   & \U                                                                                                         \\

@@

                               \end{array}\right))          . \\

@@

    (\depscohs[n][\lbrack 0,p \rbrack]:\depscohstype[n][\lbrack 0,p \rbrack])                & :                                   &
    \restrltype[n][p] (\depscohs[n][\lbrack 0,p \rbrack])                                                                                                                                               \\

@@

We can then extend $\overline{\deps}$ to build $\fulldeps$ from $\fulldepscohs$:
\begin{align*}
  \overline{\fulldeps}(\fulldepscohs[n]:\fulldepscohstype[n]) & :     \fulldepstype[n+1]                                                       \\
  \overline{\fulldeps}(\fulldepscohs[n])                      & \defeq \left(\begin{array}{l}
                                                                                 \overline{\deps}(\fulldepscohs.\depscohs[n][\lbrack 0,n \rbrack]) \\
                                                                                 \fulldepscohs.E                                                   \\
                                                                               \end{array}\right) \\
\end{align*}

@@

\begin{equation*}
  \begin{array}{llcl}
    \restrp[n+1][\lbrack 0,p-1 \rbrack]     &
    (\fulldepscohs[n]:\fulldepscohstype[n]) & :      & \restrptype[n+1][\lbrack 0,p-1 \rbrack](\fulldepscohs.\depscohs.\fulldeps[n])) \\
    \restrp[n+1]                            &
    \fulldepscohs[n]                        & \defeq & \unittype                                                                      \\
    \restrp[n+1][\lbrack 0,p \rbrack]       &
    \fulldepscohs[n]
                                            & \defeq &
    (\restrp[n+1][\lbrack 0,p-1 \rbrack](\fulldepscohs[n]),\restrp[n+1][p](\fulldepscohs[n]))                                         \\
  \end{array}
\end{equation*}

@@

\begin{align*}
  \overline{\fulldepscohs}(\fulldepscoht[n]:\fulldepscohstype[n]) & :     \fulldepscohstype[n+1]                                                       \\
  \overline{\fulldepscohs}(\fulldepscoht[n])                      & \defeq \left(\begin{array}{l}
                                                                                     \overline{\depscohs}(\fulldepscoht.\depscoht[n][\lbrack 0,n \rbrack]) \\
                                                                                     \fulldepscoht.E                                                       \\
                                                                                   \end{array}\right) \\
\end{align*}
